*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-28 Code/Manuscript_O2.Rmd
text edits

*Prochlorococcus marinus* is functionally differentiated from other cyanobacteria primarily by the absence of phycobilisomes [@chisholmProchlorococcusMarinusNov1992]. Instead, *Prochlorococcus*  uses intra-membrane Prochlorophyte chlorophyll binding (Pcb) proteins binding divinylchlorophyll *a* and *b*, and α-carotene, as major light harvesting complexes [@goerickePigmentsProchlorococcusMarinus1992]; [@larocheIndependentEvolutionProchlorophyte1996], with diversity in the number of genes encoding Pcb, expression of Pcb under changing light [@garczarekMultiplicationAntennaGenes2000; @rocapGenomeDivergenceTwo2003], and in thylakoid organization [@partenskyProchlorococcusMarinePhotosynthetic1999].  The genes encoding much of the photosynthetic apparatus are highly conserved in *Prochlorococcus* [@kettlerPatternsImplicationsGene2007], though the photosynthetic efficiency may vary among clades [@hessPhotosyntheticApparatusProchlorococcus2001; @moorePhotophysiologyMarineCyanobacterium1999].

@@

Climate change is also rapidly changing ocean chemistry.  By the end of this century, surface ocean pH is projected to decline by 0.1 to 0.4  due to projected increases in carbon dioxide concentrations [@garcia-sotoOverviewOceanClimate2021]. Moreover, substantial changes in the global water cycle, leading to extensive changes in worldwide precipitation patterns, are affecting ocean salinity levels on a global scale, and ice melts due to rising temperatures are impacting salinity levels in the Arctic and Northwest Atlantic oceans [@leeRecognizingSalinityThreats2022]. Increasing sea temperatures are also causing decreases in [O~2~] across global oceans [@matearLongtermChangesDissolved2003], particularly toward the poles [@helmObservedDecreasesOxygen2011]. Warmer ocean waters increase stratification, and decrease oxygen solubility at the surface, which in turn decreases oxygen mixing downwards by ocean currents [@garcia-sotoOverviewOceanClimate2021]. Models predict that OMZ in the Pacific and Indian Oceans are expanding [@garcia-sotoOverviewOceanClimate2021; @buseckeDivergingFatesPacific2022]. In spite of these expanding OMZ, Busecke et al. [@buseckeDivergingFatesPacific2022] project that the cores of the OMZ, where the oxygen levels are lowest, may actually contract.

We used the OceansMap Protein Portal (OPP; https://proteinportal.whoi.edu/) [@saitoDevelopmentOceanProtein2021] to analyze the distribution of proteins from clades of *P. marinus* in samples taken across a range of [O~2~] and depth, which in turn correlates to peak light at the site of sampling. We therefore analyzed the growth and physiological responses of representative strains from three clades of *P. marinus* under a matrix of [O~2~], light levels, spectral waveband ranges, and photoperiods, to approximate eco-physiological conditions representative of current and hypothetical future ocean zones. *Prochlorococcus marinus* MED4, a clade HLI strain, was isolated near the ocean surface (5m depth) of the Mediterranean Sea where [O~2~] is near saturation, light levels are high and spectral bias from full solar irradiance is minimal. *Prochlorococcus marinus* SS120, a clade LLIII strain, was isolated from the Sargasso Sea at a depth of 120 m, while *P. marinus* MIT9313, a clade LLIV strain, was isolated from the North Atlantic Gulf Stream at a depth of 135 m. At these depths, light attenuation and spectral shifts occur, resulting in low blue and green light, while [O~2~] varies from near-surface saturation levels to decreased concentrations, but does not necessarily decrease systematically with depth [@billheimerOxygenSeasonalityUtilization2021].

We then analyzed growth rates in terms of cumulative diel Photosynthetically Usable Radiation (PUR). Photosynthetic organisms absorb light energy within the Photosynthetically Active Radiation (PAR) range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. PUR represents the fraction of PAR that is absorbed by the pigments of a given photosynthetic organism [@morelAvailableUsableStored1978], taking into account the specific spectral wavebands these pigments absorb. *Prochlorococcus marinus* light-harvesting complexes, [@chisholmProchlorococcusMarinusNov1992;
@goerickePigmentsProchlorococcusMarinus1992], show an absorption maxima
of 442 nm for divinyl divinylchlorophyll *a* and 478 nm for divinyl chlorophyll b
[@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to
efficiently harvest blue light in the 400 nm
to 500 nm range [@morelAvailableUsableStored1978] prevailing in deep ocean
habitats, where only blue spectral wavelengths prevail [@holtropVibrationalModesWater2021].  In *P. marinus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993]. Given the different spectral light regimes typical of the niches of different ecotypes, expressing growth rates in terms of cumulative diel PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We also aimed to detect whether growth responses are driven simply by cumulative diel PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth.

We discuss our findings in relation to analyses of genomic sequences CITATIONS across clades of *P. marinus*, showing that differences in the presence of genes encoding oxygen-dependent enzymes, and DNA repair enzymes, can help explain differential growth responses of strains under the matrix of light and [O~2~].

@@

Collection and treatment of protein samples were performed following Saito et al. [@saitoMultipleNutrientStresses2014; @saitoNeedlesBlueSea2015]. Briefly, samples of seawater from depths of 20, 40, 50, 60, 70, 80, 90, 120, 150, 200, 250, 300, 380, 400, 500, 550, 600, and 800 m below the ocean surface were pumped through a 0.2 µm filter, preserved in RNAlater and frozen at -80°C until extraction. Proteins were extracted from the filter in an SDS-based detergent, embedded in tube gel, alkylated and reduced prior to in-gel trypsin digestion. Peptide spectra were generated using a Q-Exactive Orbitrap Mass Spectrometer, searched in the SEQUEST CITATIONXXXX and labelled with the most likely protein and species annotation from Uniprot.  Oxygen levels at the location of sampling were recorded.

@@

*In situ* measurements of Optical Density (OD) 680 nm, a proxy for cell suspension density, cell size dependent scatter and cell chlorophyll content;  and OD 720 nm, a proxy for cell suspension density and cell size dependent scatter, were recorded every 5 minutes over least 8 to 14 days,  depending on the
duration of the lag phase, if any. All data from the Multicultivator were saved
as comma separated values file  (https://github.com/FundyPhytoPhys/prochlorococcus_o2/tree/main/Data/RawData).

PAR of 180, 90 or 30 µmol photons m^-2^ s^-1^, and spectral wavebands (white LED
full spectrum, 660 nm, and 450 nm) were chosen to approximate light
levels and spectral colors spanning the vertical ocean water column, from
near-surface to the lower euphotic zone depths. Photoperiods were chosen to approximate diel cycles characteristic of current and hypothetical future
niches of *P. marinus*; 16 h represents
temperate (45°N) summer at the ocean surface; 12 h for equatorial (0°N)
ocean surface or temperate (45°N) spring and fall ocean surface or
temperate (45°N) summer at deeper ocean depths; 8 h for temperate (45°N)
winter at the surface or at temperate (45°N) spring and fall at depth
and equatorial (0°N) deep ocean depths; and 4 h for temperate (45°N)
winter or deep ocean depths during temperate (45°N) spring and fall.

## Growth rate analysis {.unnumbered}
Data files (.csv) saved from the Multicultivator software were imported into R-Studio [@R-Studio] for data management [@R-tidyverse], growth rate calculations CITATION, comparisons of model fits (get Platt CITATION), and generation of
figures using `ggplot2` [@R-ggplot2]. The chlorophyll proxy
optical density (OD680 - OD720;  ΔOD) was used to determine the chlorophyll-specific exponential growth
rate (µ, d^-1^) for each treatment combination.  We first we used a rolling mean from the RStudio `zoo` package [@R-zoo] to calculate the average ΔOD data over a 1-hour window, to lower influence of outlier points and remove data points collected during post stationary phase, if applicable. We used the Levenberg-Marquardt algorithm
[@bellaviaLevenbergMarquardtMethod2018] modification of the non-linear
least squares,  using the R package `minpack.lm`
[@R-minpack.lm], to fit a logistic
equation \@ref(eq:GrowthRate)

@@

A Generalized Additive Model (GAM) CITATION was applied to the
relation between chlorophyll-specific µ, d^-1^  to photoperiod and PAR level,  for each growth [O~2~] level, and for the blue and red wavebands for growth, for each *P.marinus* strain in this study. The gam function from the R package `mgcv`
[@R-mgcv] was used to model the growth rate with smoothing terms, to
indicate the 90, 50 and 10% quantiles for growth rate across the levels of factors. Growth rate estimates with a standard error below 30% of the fitted growth rate were included in the gam modelling. Our priority
was the effects of ecologically relevant blue light on growth trends.  We included analyses of growth responses to red light, which is not ecophysiologically relevant, but which might prove mechanistically informative [@murphyPhotoinactivationPhotosystemII2017].

@@

To estimate the Photosynthetic Usable Radiation (PUR), a proxy of incident photons absorbed by the cells, for each *P. marinus* ecotype, the imposed Photosynthetic Active
Radiation (PAR) was first determined using the reported delivery of sinusoidal diel PAR regimes by the Multicultivators, point validated using an XXXXXcitation of PAR meterXXX. An emission profile from 400 nm to 700 nm of each colored LED
light of the MCMIX-OD Multicultivator and the white LED light of the MC1000-OD Multicultivator was obtained using a Jaz spectrometer (Ocean Optics, Inc.,Dunedin, FL, USA) equipped with a fiber
optic cable, HH2 FiberOpticJmp (Part number A901073, Malvern Panalytical
Ltd, Malvern, UK). Each LED spectrum was then normalized to its emission maximum.
An *in-vivo* whole cell absorbance spectrum for each *P. marinus* strain under each spectral growth condition was obtained using the Olis 14 UV/VIS Clarity
Spectrophotometer (Olis Inc, Bogart, GA, USA) to scan across range of λ = 350 nm to 750 nm at 1 nm intervals. The path length of the internally reflective cavity of the Olis spectrophotometer was corrected to a 1 cm path length using the Javorfi correction method [@javorfiQuantitativeSpectrophotometryUsing2006] on PRO 99 media subtracted whole cell absorbance spectra. The blank-corrected whole cell absorbance spectra were normalized to the absorbance maximum of divinylchlorophyll *a* (Chl a~2~), determined for each spectra, falling between 400 nm and 460 nm.

@@

PurParPlot_cap <- glue("**Absorbed peak Photosynthetically Usable Radiation (PUR) (µmol photons m^-2^ s^-1^) vs. peak Photosynthetically Active Radiation (PAR) (µmol photons m^-2^ s^-1^).**",
                       "The correlation between PAR, plotted on the x-axis and PUR, plotted on the y-axis, are colored for each growth spectral waveband; 450 nm (blue circles), 660 nm (red circles) and white LED (black circles).",
                       "The grey dashed line represents a hypothetical one to one correlation.",
                       "**A.** is *Prochlorococcus marinus*  MED4.",
                       "**B.** is *Prochlorococcus marinus*  SS120.",
                       "**C.** is *Prochlorococcus marinus*  MIT9313.",

@@

Using the dataset Omar *et al*. (2023) [@omarAnnotationGenesEncoding2023], we filtered the dataset for Enzyme Commission Numbers (EC numbers), or Kegg Orthology Numbers (KO numbers) identified by BRENDA  [@changBRENDAELIXIRCore2021] as 'natural substrates' for O~2~ in Prochlorococcus marinus (MED4, MIT9313, SS120, and NATL2A) in order to generate a subset containing only those orthologs encoding enzymes directly mediating O~2~ metabolism. We grouped orthologs together by EC number and their KO number (KO number) and determined the occurrences of individual orthologs encoding each EC number, or KO number when EC number was not available, in a given strain. We merged the dataset with a list of enzyme Michaelis constant (K~m~) values from other organisms, as K~m~ values from Prochlorococcus were only available in the case of Ribulose bisphosphate carboxylase. XXX Possibly add the link to the full list of genes. XXXX

@@

MED4PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MED4 (High-Light (HLI) near surface clade) vs. photoperiod (h).  **",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                       "Columns show 3 levels of growth Photosynthetically Active Radiation (PAR); 30, 90 and 180 µmol photons m^-2^ s^-1^",
                    "Colors represent the actinic spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",

@@

SS120PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* SS120 (Low-Light (LLIII) deep ocean clade) vs. photoperiod (h). **",
                     "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                       "Columns show 3 levels of growth Photosynthetically Active Radiation (PAR); 30, 90 and 180 µmol photons m^-2^ s^-1^",
                    "Colors represent the actinic spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",

@@

MIT9313PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MIT9313 (Low-Light (LLIV) deep ocean clade) vs. photoperiod (h). **",
                       "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                       "Columns show 3 levels of growth Photosynthetically Active Radiation (PAR); 30, 90 and 180 µmol photons m^-2^ s^-1^",
                    "Colors represent the actinic spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",

@@

XXXX Mireille to finish Fig. caption when figure is complete. Add color ? Remove []? XXX

@@

ProchlorococcusEnzymeKms_cap <- glue("**K~m~ values for oxygen metabolizing enzymes.**" ,
                                     "The y-axis represents the log10 concentration of oxygen substrate (mM).",
                                     "The x-axis represents the oxygen metabolizing enzymes encoded in at least one of the *Prochlorococcus marinus* strains in this study.",
                                     "The *Prochlorococcus marinus* strains are indicated in rows." ,
                                     "The solid circles represent K~m~ values from literature and the asterisks represent predicted values.",
                                     "Colors represent the gene counts.",
                                     "The red shaded area denotes a K~m~ oxygen concentration range from 230 to 280 µM.",
                                     "The green shaded area denotes a K~m~ oxygen concentration range from 5 to 50 µM . ",
                                     "The blue shaded area denotes a K~m~ oxygen concentration range from 0.5 to 5 µM.  ",
                                     "The black bars show the minimum and maximum K~m~ values.",
                       "Figure was generated using a filtered subset of the annotated phytoplankton gene sequences dataset from @omarAnnotationGenesEncoding2023.",

>>>>

*Prochlorococcus marinus* is functionally differentiated from other cyanobacteria primarily by the absence of phycobilisomes [@chisholmProchlorococcusMarinusNov1992]. Instead, *Prochlorococcus*  uses intra-membrane Prochlorophyte chlorophyll binding (Pcb) proteins binding divinyl chlorophyll *a* and *b*, and α-carotene, as major light harvesting complexes [@goerickePigmentsProchlorococcusMarinus1992]; [@larocheIndependentEvolutionProchlorophyte1996], with diversity in the number of genes encoding Pcb, expression of Pcb under changing light [@garczarekMultiplicationAntennaGenes2000; @rocapGenomeDivergenceTwo2003], and in thylakoid organization [@partenskyProchlorococcusMarinePhotosynthetic1999].  The genes encoding much of the photosynthetic apparatus are highly conserved in *Prochlorococcus* [@kettlerPatternsImplicationsGene2007], though the photosynthetic efficiency may vary among clades [@hessPhotosyntheticApparatusProchlorococcus2001; @moorePhotophysiologyMarineCyanobacterium1999].

@@

Climate change is also rapidly changing ocean chemistry.  By the end of this century, surface ocean pH is projected to decline by 0.1 to 0.4  due to projected increases in carbon dioxide concentrations [@garcia-sotoOverviewOceanClimate2021]. Moreover, substantial changes in the global water cycle, leading to extensive changes in worldwide precipitation patterns, are affecting ocean salinity levels on a global scale, and ice melts due to rising temperatures are impacting salinity levels in the Arctic and Northwest Atlantic oceans [@leeRecognizingSalinityThreats2022]. Increasing sea temperatures are also causing decreases in [O~2~] across global oceans [@matearLongtermChangesDissolved2003], particularly toward the poles [@helmObservedDecreasesOxygen2011]. Warmer ocean waters increase stratification, and decrease oxygen solubility at the surface, which in turn decreases oxygen mixing downwards by ocean currents [@garcia-sotoOverviewOceanClimate2021]. Models predict that OMZ in the Pacific and Indian Oceans are expanding [@garcia-sotoOverviewOceanClimate2021; @buseckeDivergingFatesPacific2022]. In spite of these expanding OMZ, Busecke *et al*. [@buseckeDivergingFatesPacific2022] project that the cores of the OMZ, where the oxygen levels are lowest, may actually contract.

We used the OceansMap Protein Portal (OPP; https://proteinportal.whoi.edu/) [@saitoDevelopmentOceanProtein2021] to analyze the distribution of proteins from clades of *P. marinus* in samples taken across a range of [O~2~] and depth, which in turn correlates to peak light at the site of sampling. In parallel we analyzed the growth and physiological responses of representative strains from three clades of *P. marinus* under a matrix of [O~2~], light levels, spectral waveband ranges, and photoperiods, to approximate eco-physiological conditions representative of current and hypothetical future ocean zones. *Prochlorococcus marinus* MED4, a clade HLI strain, was isolated near the ocean surface (5m depth) of the Mediterranean Sea where [O~2~] is near saturation, light levels are high and spectral bias from full solar irradiance is minimal. *Prochlorococcus marinus* SS120, a clade LLIII strain, was isolated from the Sargasso Sea at a depth of 120 m, while *P. marinus* MIT9313, a clade LLIV strain, was isolated from the North Atlantic Gulf Stream at a depth of 135 m. At these depths, light attenuation and spectral shifts occur, resulting in low blue and green light, while [O~2~] varies from near-surface saturation levels to decreased concentrations, but does not necessarily decrease systematically with depth [@billheimerOxygenSeasonalityUtilization2021].

Photosynthetic organisms absorb light energy within the Photosynthetically Active Radiation (PAR) range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. Photosynthetically Usable Radiation (PUR) represents the fraction of PAR that is absorbed by the pigments of a given photosynthetic organism [@morelAvailableUsableStored1978], taking into account the specific spectral wavebands these pigments absorb. *Prochlorococcus marinus* light-harvesting complexes, [@chisholmProchlorococcusMarinusNov1992;
@goerickePigmentsProchlorococcusMarinus1992], show an absorption maxima
of 442 nm for divinyl chlorophyll *a* and 478 nm for divinyl chlorophyll *b*
[@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to
efficiently harvest blue light in the 400 nm
to 500 nm range [@morelAvailableUsableStored1978] prevailing in deep ocean
habitats, where only blue spectral wavelengths prevail [@holtropVibrationalModesWater2021].  In *P. marinus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993]. Given the different spectral light regimes typical of the niches of different ecotypes, expressing growth rates in terms of cumulative diel PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We also aimed to detect whether growth responses are driven simply by cumulative diel PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth. We therefore analyzed growth rates in terms of cumulative diel PUR.

We discuss our findings in relation to analyses of genomic sequences CITATIONS across clades of *P. marinus*, showing that differences in the presence of genes encoding oxygen-dependent enzymes, and DNA repair enzymes, can help explain differential growth responses of strains under the matrix of light and [O~2~] conditions.

@@

Collection and treatment of protein samples were performed by Saito *et al*. [@saitoMultipleNutrientStresses2014; @saitoNeedlesBlueSea2015]. Briefly, samples of seawater from depths of 20, 40, 50, 60, 70, 80, 90, 120, 150, 200, 250, 300, 380, 400, 500, 550, 600, and 800 m below the ocean surface were pumped through a 0.2 µm filter, preserved in RNAlater and frozen at -80°C until extraction. Proteins were extracted from the filter in an SDS-based detergent, embedded in tube gel, alkylated and reduced prior to in-gel trypsin digestion. Peptide spectra were generated using a Q-Exactive Orbitrap Mass Spectrometer, searched in the SEQUEST CITATIONXXXX and labelled with the most likely protein and species annotation from Uniprot.  Oxygen levels at the location of sampling were recorded.

@@

*In situ* measurements of Optical Density (OD) 680 nm, a proxy for cell suspension density, cell size dependent scatter and cell chlorophyll content;  and OD 720 nm, a proxy for cell suspension density and cell size dependent scatter, were recorded every 5 minutes over least 8 to 14 days,  depending on the
duration of the lag phase, if any. All data from the Multicultivator were saved
as comma separated values files  (https://github.com/FundyPhytoPhys/prochlorococcus_o2/tree/main/Data/RawData).

PAR of 180, 90 or 30 µmol photons m^-2^ s^-1^, and spectral wavebands (white LED
full spectrum, 660 nm, and 450 nm) were chosen to approximate light
levels and spectral colours spanning the vertical ocean water column, from
near-surface to the lower euphotic zone depths. Photoperiods were chosen to approximate diel cycles characteristic of current and hypothetical future
niches of *P. marinus*; 16 h represents
temperate (45°N) summer at the ocean surface; 12 h for equatorial (0°N)
ocean surface or temperate (45°N) spring and fall ocean surface or
temperate (45°N) summer at deeper ocean depths; 8 h for temperate (45°N)
winter at the surface or at temperate (45°N) spring and fall at depth
and equatorial (0°N) deep ocean depths; and 4 h for temperate (45°N)
winter or deep ocean depths during temperate (45°N) spring and fall.

## Growth rate analysis {.unnumbered}
Data files (.csv) saved from the Multicultivator software were imported into R-Studio [@R-Studio] for data management [@R-tidyverse], growth rate calculations CITATION, comparisons of model fits (get Platt CITATION), and generation of
figures using `ggplot2` [@R-ggplot2]. The chlorophyll proxy
optical density (OD680 - OD720;  ΔOD) was used to determine the chlorophyll-specific exponential growth
rate (µ, d^-1^) for each treatment combination.  We first used a rolling mean from the RStudio `zoo` package [@R-zoo] to calculate the average ΔOD data over a 1-hour window to lower influence of outlier points and remove data points collected during post stationary phase, if applicable. We used the Levenberg-Marquardt algorithm
[@bellaviaLevenbergMarquardtMethod2018] modification of the non-linear
least squares,  using the R package `minpack.lm`
[@R-minpack.lm], to fit a logistic
equation \@ref(eq:GrowthRate)

@@

A Generalized Additive Model (GAM) CITATION was applied to the
relation between chlorophyll-specific µ, d^-1^  to photoperiod and PAR level,  for each growth [O~2~] level, and for the blue and red wavebands for growth, for each *P.marinus* strain in this study. The gam function from the R package `mgcv`
[@R-mgcv] was used to model the growth rate with smoothing terms, to
indicate the 90, 50 and 10% quantiles for growth rate across the levels of factors. Growth rate estimates with a standard error below 30% of the fitted growth rate were included in the GAM modelling. Our priority
was the effects of ecologically relevant blue light on growth trends.  We included analyses of growth responses to red light, which is not ecophysiologically relevant, but which might prove mechanistically informative [@murphyPhotoinactivationPhotosystemII2017].

@@

To estimate the Photosynthetic Usable Radiation (PUR), a proxy of incident photons absorbed by the cells, for each *P. marinus* ecotype, the imposed Photosynthetic Active
Radiation (PAR) was first determined using the reported delivery of sinusoidal diel PAR regimes by the Multicultivators, point validated using an XXXXXcitation of PAR meterXXX. An emission profile from 400 nm to 700 nm of each coloured LED
light of the MCMIX-OD Multicultivator and the white LED light of the MC1000-OD Multicultivator was obtained using a Jaz spectrometer (Ocean Optics, Inc.,Dunedin, FL, USA) equipped with a fiber
optic cable, HH2 FiberOpticJmp (Part number A901073, Malvern Panalytical
Ltd, Malvern, UK). Each LED spectrum was then normalized to its emission maximum.
An *in-vivo* whole cell absorbance spectrum for each *P. marinus* strain under each spectral growth condition was obtained using the Olis 14 UV/VIS Clarity
Spectrophotometer (Olis Inc, Bogart, GA, USA) to scan across range of λ = 350 nm to 750 nm at 1 nm intervals. The path length of the internally reflective cavity of the Olis spectrophotometer was corrected to a 1 cm path length using the Javorfi correction method [@javorfiQuantitativeSpectrophotometryUsing2006] on PRO 99 media subtracted whole cell absorbance spectra. The blank-corrected whole cell absorbance spectra were normalized to the absorbance maximum of divinyl chlorophyll *a* (Chl a~2~), determined for each spectra, falling between 400 nm and 460 nm.

@@

PurParPlot_cap <- glue("**Absorbed peak Photosynthetically Usable Radiation (PUR) (µmol photons m^-2^ s^-1^) vs. peak Photosynthetically Active Radiation (PAR) (µmol photons m^-2^ s^-1^).**",
                       "The correlation between PAR, plotted on the x-axis and PUR, plotted on the y-axis, are coloured for each growth spectral waveband; 450 nm (blue circles), 660 nm (red circles) and white LED (black circles).",
                       "The grey dashed line represents a hypothetical one to one correlation.",
                       "**A.** is *Prochlorococcus marinus*  MED4.",
                       "**B.** is *Prochlorococcus marinus*  SS120.",
                       "**C.** is *Prochlorococcus marinus*  MIT9313.",

@@

Using the dataset Omar *et al*. [@omarAnnotationGenesEncoding2023], we filtered the dataset for Enzyme Commission Numbers (EC numbers), or Kegg Orthology Numbers (KO numbers) identified by BRENDA  [@changBRENDAELIXIRCore2021] as 'natural substrates' for O~2~ in Prochlorococcus marinus (MED4, MIT9313, SS120, and NATL2A) in order to generate a subset containing only those orthologs encoding enzymes directly mediating O~2~ metabolism. We grouped orthologs together by EC number and their KO number (KO number) and determined the occurrences of individual orthologs encoding each EC number, or KO number when EC number was not available, in a given strain. We merged the dataset with a list of enzyme Michaelis constant (K~m~) values from other organisms, as K~m~ values from Prochlorococcus were only available in the case of Ribulose bisphosphate carboxylase. XXX Possibly add the link to the full list of genes. XXXX

@@

MED4PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MED4 (High-Light (HLI) near surface clade) vs. photoperiod (h).  **",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                       "Columns show 3 levels of growth Photosynthetically Active Radiation (PAR); 30, 90 and 180 µmol photons m^-2^ s^-1^",
                    "colours represent the actinic spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",

@@

SS120PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* SS120 (Low-Light (LLIII) deep ocean clade) vs. photoperiod (h). **",
                     "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                       "Columns show 3 levels of growth Photosynthetically Active Radiation (PAR); 30, 90 and 180 µmol photons m^-2^ s^-1^",
                    "colours represent the actinic spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",

@@

MIT9313PAR_cap <- glue("**Chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MIT9313 (Low-Light (LLIV) deep ocean clade) vs. photoperiod (h). **",
                       "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                       "Columns show 3 levels of growth Photosynthetically Active Radiation (PAR); 30, 90 and 180 µmol photons m^-2^ s^-1^",
                    "colours represent the actinic spectral waveband (nm).",
                    "Large circles show mean or single determinations of growth rate from logistic curve fits; small circles show values for replicate determinations, if any: replicates often fall with larger circles.",

@@

XXXX Mireille to finish Fig. caption when figure is complete. Add colour ? Remove []? XXX

@@

ProchlorococcusEnzymeKms_cap <- glue("**K~m~ values for oxygen metabolizing enzymes.**" ,
                                     "The y-axis represents the log10 concentration of oxygen substrate (mM).",
                                     "The x-axis represents the oxygen metabolizing enzymes encoded in at least one of the *Prochlorococcus marinus* strains in this study.",
                                     "The *Prochlorococcus marinus* strains are indicated in rows." ,
                                     "The solid circles represent K~m~ values from literature and the asterisks represent predicted values.",
                                     "colours represent the gene counts.",
                                     "The red shaded area denotes a K~m~ oxygen concentration range from 230 to 280 µM.",
                                     "The green shaded area denotes a K~m~ oxygen concentration range from 5 to 50 µM . ",
                                     "The blue shaded area denotes a K~m~ oxygen concentration range from 0.5 to 5 µM.  ",
                                     "The black bars show the minimum and maximum K~m~ values.",
                       "Figure was generated using a filtered subset of the annotated phytoplankton gene sequences dataset from @omarAnnotationGenesEncoding2023.",
